\documentclass[dvipsnames]{article}
\usepackage{titlesec}

\titleformat{\section}
  {\normalfont\Large\bfseries}
  {Unit \thesection:}{1em}{}

\renewcommand{\thesection}{\arabic{section}}

\usepackage{marvosym}
\input{global}
\input{notes}
\input{physics-glossary}
\usepackage{figchild}
\makenoidxglossaries

\usepackage{circuitikz}
\ctikzset{bipoles/battery1/height=0.5}
\ctikzset{bipoles/battery1/width=0.25}
\ctikzset{bipoles/resistor/height=0.15}
\ctikzset{bipoles/resistor/width=0.4}
\ctikzset{bipoles/cuteswitch/thickness=0.3,bipoles/cuteswitch/shape=circ}
\ctikzset{bipoles/bulb/height=0.4,bipoles/bulb/width=0.4}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%...Hide or show unit contents below:            
\newif\ifShowUnitI                               
\newif\ifShowUnitII                                      
\newif\ifShowUnitIII                                     
\newif\ifShowUnitIV                                      
\newif\ifShowUnitV                                       
\newif\ifShowUnitVI                                      
\newif\ifShowUnitVII                                     
\newif\ifShowUnitVIII                                    
\newif\ifShowUnitIX                                      
\newif\ifShowUnitX                                       
\newif\ifShowUnitXI                                      
\newif\ifShowUnitXII
\newif\ifShowUnitXIII
\newif\ifShowUnitXIV
\ShowUnitIfalse                                  
\ShowUnitIIfalse                                 
\ShowUnitIIIfalse                                
\ShowUnitIVfalse                                 
\ShowUnitVfalse                                  
\ShowUnitVItrue                                 
\ShowUnitVIItrue                              
\ShowUnitVIIIfalse                               
\ShowUnitIXfalse                                 
\ShowUnitXfalse                                  
\ShowUnitXIfalse                                 
\ShowUnitXIItrue  
% \ShowUnitXIIItrue
% \ShowUnitXIVtrue  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\title{Physics}
\author{The Curriculum at Cypress Springs High School}
\date{Updated on \today}

\begin{document}

\maketitle
\tableofcontents
\clearpage

\section*{Mission Statement}

The following are goals we value Cypress Springs High School:

\begin{enumerate}[itemsep=0pt]
    \item To provide students with foundational knowledge in physics.
\end{enumerate}

We value a curriculum that

\begin{itemize}[itemsep=0pt]
    \item is clear
    \item is coherent
    \item is based on knowledge and fundamental principles
    \item equips students with basic vocabulary terms all physics students should know
    \item has a robust sequence that parallels the curricular development of the field's best textbooks
    \item connects students to history and to notable figures like Newton and Faraday
    \item strikes a balance between abstract and concrete principles; between mathematical and conceptual problems; between words and images
    \item aims to reduce cognitive overload in students
    \item exposes students to geometric and graphical figures of the highest quality
\end{itemize}


\section{Constant Velocity}

\subsection*{Vocabulary}

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item Define \textbf{\gls{magnitude}} as \glsdesc{magnitude}.
    \item Define \textbf{\gls{scalar}} as \glsdesc{scalar}.
    \item Define \textbf{\gls{vector}} as \glsdesc{vector}.
    \item Define \textbf{\gls{component}} as \glsdesc{component}.
    \item Define \textbf{\gls{tail}} as \glsdesc{tail}.
    \item Define \textbf{\gls{head}} as \glsdesc{head}.
    \item Define \textbf{\gls{head-to-tail method}} as \glsdesc{head-to-tail method}.
    \item Define \textbf{\gls{vector sum}} as \glsdesc{vector sum}.
    \item Define \textbf{\gls{position}} ($x$) as \glsdesc{position}.
    \item Define \textbf{\gls{motion}} as \glsdesc{motion}.
    \item Define \textbf{\gls{reference frame}} as \glsdesc{reference frame}.
    \item Define \textbf{\gls{one-dimensional motion}} as \glsdesc{one-dimensional motion}.
    \item Define \textbf{\gls{displacement}} ($\Delta x$) as \glsdesc{displacement}.
    \item Define \textbf{\gls{distance}} ($d$) as \glsdesc{distance}.
    \item Define \textbf{\gls{position vs. time graph}} as \glsdesc{position vs. time graph}.
    \item Define \textbf{\gls{speed}} as \glsdesc{speed}.
    \item Define \textbf{\gls{average speed}} as \glsdesc{average speed}.
    \item Define \textbf{\gls{velocity1}} as \glsdesc{velocity1}.
    \item Define \textbf{\gls{velocity2}} ($v$) as \glsdesc{velocity2}.
    \item Define \textbf{\gls{average velocity}} ($\bar{v}$) as \glsdesc{average velocity}.
    \item Define \textbf{\gls{velocity vs. time graph}} as \glsdesc{velocity vs. time graph}.
\end{itemize}

\bigskip

\subsection*{Fundamental Principles}

% Students must articulate the following physical rules governing one-dimensional motion:

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Know} that a vector is graphically represented as an arrow.
    \item \textbf{Know} that a vector is positive if it is directed to the right, and negative if it is directed to the left.
    \item \textbf{Know} that position, velocity, and displacement are vectors.
    \item \textbf{Know} that distance, speed, and time are scalars.
    \item \textbf{Know} that positive direction for position or velocity is to the right on a horizontal position axis, and negative direction is to the left.
    \begin{center}
        \begin{tikzpicture}[y=1cm,x=1cm]
            \foreach \x in {-5,-4,...,5} {
                \draw[ultra thin] (\x,-0.09) -- (\x,0.09) node[below=0.2] {$\x$};
            }
            \draw[->] (-5,0) -- (5,0) node[below=0.6,pos=0.5] {Position, $x$ (m)};
            \draw[->,ultra thick] (-1.5,0.4) -- (1.5,0.4) node[above,pos=0.5] {positive direction};
            \draw[->,ultra thick] (1.5,1.2) -- (-1.5,1.2) node[above,pos=0.5] {negative direction};
            % \draw[thick,<-] (0,0.15) -- ++(0,0.5);
        \end{tikzpicture}
    \end{center}
    \item \textbf{Know} that an object's position vector starts at the origin (at zero) and ends at the object's location on the position axis.
    \item \textbf{Understand} that an object is in motion if its position is changing.

    \begin{center}
        \begin{tikzpicture}[y=1cm,x=1cm]
            \foreach \x in {-5,-4,...,5} {
                \draw[ultra thin] (\x,-0.09) -- (\x,0.09) node[below=0.2] {$\x$};
            }
            \draw[->] (-5,0) -- (5,0) node[below=0.6,pos=0.5] {Position (m)};
            \node[above,black!20] at (0,0) {\Strichmaxerl[2]};
            \node[above,black!40] at (1.5,0) {\Strichmaxerl[2]};
            \node[above] at (3,0) {\Strichmaxerl[2]};
            \draw[black!80] (2.5,0.4) -- ++(0.3,0) (2.4,0.3) -- ++(0.4,0) (2.5,0.2) -- ++(0.3,0);
            \draw[thick,->] (2.8,0.8) -- ++(0.6,0) node[above,pos=0.5] {$\vec{v}$};
        \end{tikzpicture}
    \end{center}
    \item \textbf{Understand} that constant velocity implies that displacement is constant across equal time intervals.
    \item \textbf{Understand} that a vector's direction is indicated by a positive ($+$) or negative ($-$) sign\\(e.g., $\vec{v} = \SI{-3}{m/s}$,  $\vec{x} = \SI{8}{m}$).
    
    \item \textbf{Understand} that an object is moving away from the origin if its position and velocity vectors point in the same directions (both positive or both negative).

    \begin{center}
        \begin{tikzpicture}[y=1cm,x=1cm]
            \foreach \x in {-5,-4,...,5} {
                \draw[ultra thin] (\x,-0.09) -- (\x,0.09) node[below=0.2] {$\x$};
            }
            \draw[->] (-5,0) -- (5,0) node[below=0.6,pos=0.5] {Position (m)};
            \draw[thick,->] (0,0.3) -- ++(-1.9,0) node[above,pos=0.5] {$\vec{x}$};
            \node[above] at (-2,0) {\Strichmaxerl[2]};
            \draw[very thick,->] (-2,0.7) -- ++(-1,0) node[above,pos=0.5] {$\vec{v}$};
        \end{tikzpicture}
    \end{center}
    \item \textbf{Understand} that an object is moving towards the origin if its position and velocity vectors point in opposite directions (one is positive, the other is negative).
    % \begin{center}
    %     \begin{tikzpicture}[y=1cm,x=1cm]
    %         \foreach \x in {-5,-4,...,5} {
    %             \draw[ultra thin] (\x,-0.09) -- (\x,0.09) node[below=0.2] {$\x$};
    %         }
    %         \draw[->] (-5,0) -- (5,0) node[below=0.6,pos=0.5] {Position (m)};
    %         \draw[thick,->] (0,0.3) -- ++(2.9,0) node[above,pos=0.5] {$\vec{x}$};
    %         \node[above] at (3,0) {\Strichmaxerl[2]};
    %         \draw[very thick,->] (3,0.7) -- ++(-1,0) node[above,pos=0.5] {$\vec{v}$};
    %     \end{tikzpicture}
    % \end{center}
    \item \textbf{Distinguish} between distance (a scalar) and displacement (a vector).
    \item \textbf{Understand} that an object can have a position of zero but still be moving (e.g., at the origin of the reference frame).
    \item \textbf{Distinguish} between speed (a scalar) and velocity (a vector).
\end{itemize}

\bigskip

\subsection*{Mathematical Formulas}

% Students must memorize the four core kinematic equations for constant acceleration and identify which variables are present and missing in each:



\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Know} the variables used to represent physical quantities, including knowing that $x$ stands position, $\Delta x$ stands for displacement, $t$ stands for time, $v$ stands for velocity.
    \item \textbf{Know} that the subscripts $i$ and $f$ are used to denote initial and final variables, respectively (e.g.,\\$v_i$ = initial velocity, $x_f$ = final position, etc.).
    \item \textbf{Know} that the base SI unit of position is the meter (m), the base SI unit of time is the second (s), and the base SI unit of velocity is the meter per second (m/s).
    \item \textbf{Know} that average speed is given by distance divided by time.
    \item \textbf{Know} that average velocity is given by $\displaystyle \bar{v} = \frac{\Delta x}{\Delta t}$.
    \item \textbf{Know} that slope is rise over run.
\end{itemize}

\bigskip



\subsection*{Learning Goals}

% Students must demonstrate mastery of the concepts and equations by performing the following specific tasks:

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Calculate} a vector sum for two or more vectors (1) by the head-to-tail method and (2) arithmetically.
    \item \textbf{Identify} the origin in a position reference frame.
    \item \textbf{Draw} the position vector for an object on a position axis.
    \item \textbf{Draw} a motion map for an object undergoing constant velocity.
    \begin{center}
    \begin{tikzpicture}
        \foreach \x in {-5,-4,...,5} {
            \fill (\x,0) circle (0.07);
        } 
    \end{tikzpicture}
    \end{center}
    \item \textbf{Calculate} displacement given initial and final positions.
    \item \textbf{Predict} data in a table containing position and time values.
    % \item \textbf{Combine} one-dimensional vectors using both graphical vector addition.
    \item \textbf{Apply} correct sign (positive or negative) appropriately on vector quantities based on the established reference frame.
    % \item \textbf{Extract} given values from a word problem and correctly assign them to the variables $v_i$, $v_f$, $a$, $t$, and $\Delta x$.
    % \item \textbf{Identify} the single missing variable requested by the problem and select the correct kinematic equation that contains the knowns and the unknown.
    % \item \textbf{Isolate} the unknown variable algebraically before plugging in numerical values.
    \item \textbf{Find} average velocity ($\bar{v}$) on a position vs.\,time graph by calculating the slope using two points on the line. 
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \draw[->] (0,0) -- (0,1.2) node[left,pos=0.5] {$x$};
        \draw[->] (0,0) -- (1.2,0) node[below,pos=0.5] {$t$};
        \draw[thick] (0,0) -- ($(0,0)!1.1!(1,0.8)$);
        \fill (0.4,0.32) circle (0.03);
        \fill (0.9,0.72) circle (0.03);
        \draw (0.4,0.32) -- (0.4,0.72) node[pos=0.5,left] {$\Delta x$} -- (0.9,0.72) node[above,pos=0.5] {$\Delta t$};
        \node[overlay] at (1.1,0.4) {$\displaystyle \bar{v} = \frac{\Delta x}{\Delta x}$};
    \end{tikzpicture}
    \end{center}
    \item \textbf{Find} displacement on a velocity vs.\,time graph by calculating the rectangular area bounded by the line.
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \fill[black!10] (0,0.5) -- (1,0.5) -- (1,0) -- (0,0) -- cycle;
        \node at (1.1,0.25) {$\Delta x = \text{base} \times \text{height}$};
        \draw[->] (0,0) -- (0,1) node[left,pos=0.8] {$v$};
        \draw[->] (0,0) -- (1.2,0) node[below,pos=0.5] {$t$};
        \draw[thick] (0,0.5) -- (1,0.5);
    \end{tikzpicture}
    \end{center}
\end{itemize}

\subsection*{Assessment}

\begin{enumerate}[topsep=-3pt,itemsep=0pt]
    \item Find the vector sum of all vectors shown below. Each square represents 1 unit of magnitude.
    \begin{center}
        %...PRO TIP: Use a random number generator to generate random vectors between -10 and 10
        \begin{tikzpicture}[x=5mm,y=5mm]
            \draw[step=1,lightgray] (0,0) grid (10,4);
            \draw[->,line width=0.6mm,-{Triangle}] (5,4) -- ++(-3,0);
            \draw[->,line width=0.6mm,-{Triangle}] (9,3) -- ++(-8,0);
            \draw[->,line width=0.6mm,-{Triangle}] (2,2) -- ++(6,0);
            \draw[->,line width=0.6mm,-{Triangle}] (6,1) -- ++(-4,0);
            \draw[->,line width=0.6mm,-{Triangle}] (0,0) -- ++(5,0);
        \end{tikzpicture}
    \end{center}

    \item Charlie moves from point A to point B. Calculate (a) the distance traveled and (b) the displacement. (c) Compare and contrast the distance and displacement.

    \begin{center}
        \begin{tikzpicture}[y=1cm,x=1cm]
            \foreach \x in {-5,-4,...,5} {
                \draw[ultra thin] (\x,-0.09) -- (\x,0.09) node[below=0.2] {$\x$};
            }
            \draw[->] (-5,0) -- (5,0) node[below=0.6,pos=0.5] {Position (m)};
            \fill (3,0) circle (0.1) node[above=0.1] {A};
            \fill (-4,0) circle (0.1) node[above=0.1] {B};
        \end{tikzpicture}
    \end{center}

    \item Dana moves from point A to point B in 10 seconds. Calculate (a) the average speed and (b) the average velocity. (c) Compare and contrast the average speed and average velocity.

    \begin{center}
        \begin{tikzpicture}[y=3mm,x=3mm]
            \foreach \x in {-20,-15,...,20} {
                \draw[ultra thin] (\x,-1mm) -- (\x,1mm) node[below=3mm] {$\x$};
            }
            \draw[->] (-20,0) -- (20,0) node[below=7mm,pos=0.5] {Position (m)};
            \fill (10,0) circle (1mm) node[above=1mm] {A};
            \fill (-20,0) circle (1mm) node[above=1mm] {B};
        \end{tikzpicture}
    \end{center}
    
    \item Predict the last row of the data table below, assuming constant velocity.
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \thead{Position (m)} & \thead{Time (s)} \\ \hline
            0 & 0 \\ \hline
            3 & 1 \\ \hline
            6 & 2 \\ \hline
            9 & 3 \\ \hline
              & 4 \\ \hline
        \end{tabular}
    \end{center}
\end{enumerate}

\clearpage

\section{Acceleration}

\subsection*{Vocabulary}

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item Define \textbf{\gls{acceleration}} as \glsdesc{acceleration}; or as \glsdesc{acceleration2}.
    \item Define \textbf{average acceleration} as \glsdesc{average acceleration}.
    \item Define \textbf{acceleration vs. time graph} as \glsdesc{acceleration vs. time graph}.
\end{itemize}

\medskip

\subsection*{Fundamental Principles}

% Students must articulate the following physical rules governing one-dimensional motion:

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Recall} that $x$ stands for position, $v$ stands for velocity, and $t$ stands for time.
    \item \textbf{Recall} that the position axis conventionally determines the sign (direction) of velocity and acceleration vectors.
    \item \textbf{Know} that an object is accelerating if its velocity is changing.
    \item \textbf{Recognize} that this unit (and course) deals only with constant (uniform) acceleration.
    \item \textbf{Know} that acceleration is a vector, so that the sign ($+$ or $-$) of acceleration indicates the direction.
    \item \textbf{Understand} that an object is speeding up if its velocity and acceleration vectors point in the same direction (both positive or both negative).
    \item \textbf{Understand} that an object is slowing down if its velocity and acceleration vectors point in opposite directions (one is positive, the other is negative).
    \item \textbf{Understand} that constant velocity implies that acceleration is exactly zero.
    \item \textbf{Understand} that an object can have a velocity of zero but still be accelerating (e.g., at the peak of a thrown ball's trajectory).
    \item \textbf{Understand} that, if an object is accelerating, the velocity vs.\,time graph has a slope that is constant.
    \item \textbf{Know} that the slope of a velocity vs.\,time graph is acceleration.
    \item \textbf{Understand} that, if an object is accelerating, the position vs.\,time graph has a slope that is changing.
    \item \textbf{Know} that the area bounded by a line in a velocity vs.\,time graph is displacement.
    \item \textbf{Know} that the area bounded by a line in an acceleration vs.\,time graph is change in velocity.
\end{itemize}

\subsection*{Mathematical Formulas}

\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Know} variables used to represent physical quantities, including: $a$ = acceleration; $\Delta v$ = change in velocity; $\Delta t$ = change in time.
    \item \textbf{Know} that the base SI unit of acceleration is the meter per second per second (m/s/s), or the meter per second squared ($\mathrm{m/s^2}$).
    \item \textbf{Know} that average acceleration is given by $\displaystyle a = \frac{\Delta v}{\Delta t}$.
    \item \textbf{Know} that the area of a triangle is given by $\frac{1}{2}\times \text{base} \times \text{height}$.
\end{itemize}



\subsection*{Learning Goals}

\begin{enumerate}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Justify} whether or not an object is accelerating given a data table of velocity and time.
    \item \textbf{Justify} whether or not an object is accelerating given a data table of position and time.
    \item \textbf{Draw} a motion map for an object undergoing positive or negative acceleration.
    \item \textbf{Indicate} whether an object is speeding up or slowing down from a motion map.
    \item \textbf{Identify} whether or not an object is accelerating from a velocity vs.\,time graph.
    \item \textbf{Identify} the direction of acceleration from a velocity vs.\,time graph.
    \item \textbf{Identify} the sign (positive or negative) of the slope of a point in a position vs.\,time graph, and associate it with the direction of the object's velocity.
    % \item \textbf{Identify} whether an object is speeding up or slowing down from a positive vs.\,time graph.
    \item \textbf{Estimate} changes (increasing or decreasing) in slope across time in a position vs.\,time, to \textbf{determine} (a) whether an object is speeding up or slowing down and (b) the sign of acceleration.
    \item \textbf{Pair} position vs.\,time and velocity vs.\,time graphs of a common object undergoing acceleration.
    \item \textbf{Calculate} change in velocity given a particle's initial and final velocities.
    \item \textbf{Calculate} average acceleration given a particle's initial and final velocities and the time of travel.
    \item \textbf{Calculate} average acceleration by finding the slope in a velocity vs.\,time graph.
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \draw[->] (0,0) -- (0,1.2) node[left,pos=0.5] {$v$};
        \draw[->] (0,0) -- (1.2,0) node[below,pos=0.5] {$t$};
        \draw[thick] (0,0) -- ($(0,0)!1.1!(1,0.8)$);
        \fill (0.4,0.32) circle (0.03);
        \fill (0.9,0.72) circle (0.03);
        \draw (0.4,0.32) -- (0.4,0.72) node[pos=0.5,left] {$\Delta v$} -- (0.9,0.72) node[above,pos=0.5] {$\Delta t$};
        \node[overlay] at (1.1,0.4) {$\displaystyle \bar{a} = \frac{\Delta v}{\Delta x}$};
    \end{tikzpicture}
    \end{center}
    \item \textbf{Calculate} displacement of an accelerating object by finding the triangular area bounded by a line in a velocity vs.\,time graph.
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \fill[black!10] (0,0) -- (0.8,0) -- ($(0,0)!0.8!(1,0.8)$) -- cycle;
        \draw[->] (0,0) -- (0,1) node[left,pos=0.5] {$v$};
        \draw[->] (0,0) -- (1,0) node[below,pos=0.5] {$t$};
        \draw[thick] (0,0) -- ($(0,0)!0.8!(1,0.8)$);
        \node at (1.4,0.2) {$\Delta x = \frac{1}{2} \times \text{base} \times \text{height}$};
    \end{tikzpicture}
    \end{center}
    \item \textbf{Calculate} change in velocity ($\Delta v$) on an acceleration vs.\,time graph by finding the rectangular area bounded by the line.
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \fill[black!10] (0,0.5) -- (1,0.5) -- (1,0) -- (0,0) -- cycle;
        \node at (1.1,0.25) {$\Delta v = \text{base} \times \text{height}$};
        \draw[->] (0,0) -- (0,1) node[left,pos=0.8] {$a$};
        \draw[->] (0,0) -- (1.2,0) node[below,pos=0.5] {$t$};
        \draw[thick] (0,0.5) -- (1,0.5);
    \end{tikzpicture}
    \end{center}
\end{enumerate}

\subsection*{Assessment}

\begin{enumerate}[itemsep=0pt,topsep=-3pt]
    \item An object's motion is summarized in the table below. Is the object accelerating?
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \thead{Velocity (m/s)} & \thead{Time (s)} \\ \hline
            0 & 0 \\ \hline
            2 & 1 \\ \hline
            4 & 2 \\ \hline
            6 & 3 \\ \hline
            8 & 4 \\ \hline
        \end{tabular}
    \end{center}
    
    \item Give the data below, is the object accelerating?
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \thead{Position (m)} & \thead{Time (s)} \\ \hline
            4 & 0 \\ \hline
            8 & 1 \\ \hline
            12 & 2 \\ \hline
            16 & 3 \\ \hline
            20 & 4 \\ \hline
        \end{tabular}
    \end{center}
    
    \item True or false? The graph below represents accelerating motion.
    \begin{center}
    \begin{tikzpicture}
        \begin{axis}[width=5cm,height=4cm,
            axis lines=left,
            xlabel = {Time (s)},
            ylabel = {Velocity (m/s)},
            ymin=0, ymax=8,
            xmin=0, xmax=8,
            ytick={0,2,...,8},
            xtick={0,2,...,8},
            grid=both,
            minor x tick num = 1,
            minor y tick num = 1,
        ]
            \addplot[very thick,black,domain=0:8] {5};
            \end{axis}
    \end{tikzpicture}
    \end{center}

    \item The motion map below depicts an object moving with positive velocity. Is the object speeding up, slowing down, or moving at constant speed?

    \begin{center}
        \begin{tikzpicture}[y=15cm,x=15cm]
            \foreach \x in {0,0.1,...,1} {
                \fill (-\x*\x,0) circle (1mm);
            }
            \draw[thick,->] (-0.81,0.05) -- ++(0.07,0) node[right] {$+x$};
        \end{tikzpicture}
    \end{center}
    
    \item What is the direction of the object's velocity as depicted in the figure below?
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=1.5cm]
        \draw[->] (0,-1) -- (0,1) node[pos=0.8,left] {$x$} node[pos=0.5,left] {\footnotesize $0$};
        \draw[->] (0,0) -- (1,0) node[right] {$t$};
        \draw[very thick,domain=0:0.7,smooth] plot(\x,{-1.1*(\x-0.8)^2});
    \end{tikzpicture}
    \end{center}
    
    \item Does the graph below represent an object speeding up or slowing down?
    \begin{center}
    \begin{tikzpicture}[x=2cm,y=2cm]
        \draw[->] (0,0) -- (0,1) node[pos=0.5,left] {$x$} node[pos=0,shift={(-0.05,-0.05)}] {\footnotesize $0$};
        \draw[->] (0,0) -- (1,0) node[pos=0.5,below] {$t$};
        \draw[very thick,domain=0:0.8,smooth] plot(\x,{(\x-0.8)^2});
    \end{tikzpicture}
    \end{center}

    \item What is the direction of the object's acceleration?
    \begin{center}
    \begin{tikzpicture}[x=2.5cm,y=1.5cm]
        \draw[->] (0,-1) -- (0,1) node[pos=0.8,left] {$x$} node[pos=0.5,left] {\footnotesize $0$};
        \draw[->] (0,0) -- (1,0) node[right] {$t$};
        \draw[very thick,domain=0:1,smooth] plot(\x,{6*(\x-0.5)^2 - 0.8});
    \end{tikzpicture}
    \end{center}

    \item An accelerating particle starts at a speed of \SI{9}{m/s} in the positive $x$ direction. Some time later, it moves at \SI{4}{m/s} in the negative direction. Calculate the change in velocity.

    \item An accelerating particle starts at a speed of \SI{3}{m/s} in the negative $x$ direction. After 4 seconds, it moves at \SI{5}{m/s} in the positive direction. Calculate the average acceleration.
    
    \item Calculate acceleration from the velocity vs.\,time graph below.
    \begin{center}
    \begin{tikzpicture}
    \begin{axis}[width=5.2cm,height=4.2cm,
        axis lines=left,
        % axis x line=center,
        xlabel = {Time (s)},
        ylabel = {Velocity (m/s)},
        ymin=0, ymax=14,
        xmin=0, xmax=10,
        ytick={0,2,...,14},
        xtick={0,1,...,10},
        ymajorgrids=true,
        xmajorgrids=true,
        ]
        \addplot[very thick,black,domain=0:10] {-3/2*\x+18};
        \end{axis}
    \end{tikzpicture}
    \end{center}

    \item (a) Calculate the object's displacement across the entire time interval shown below. (b) Calculate the average acceleration.
    \begin{center}
    \begin{tikzpicture}
    \begin{axis}[width=5.2cm,height=4.2cm,
        axis lines=left,
        xlabel = {Time (s)},
        ylabel = {Velocity (m/s)},
        ymin=0, ymax=8,
        xmin=0, xmax=8,
        ytick={0,2,...,8},
        xtick={0,2,...,8},
        minor y tick num=1,
        minor x tick num=1,
        grid=both,
        ]
        \addplot[very thick,domain=0:8] {6/8*x};
        \end{axis}
    \end{tikzpicture}
    \end{center}

    \item What is the change in the object's velocity?
    \begin{center}
    \begin{tikzpicture}
        \begin{axis}[width=5.5cm,height=4.5cm,
            axis lines=center,
            xlabel = {Time (s)},
            ylabel = {Acceleration ($\mathrm{m/s^2}$)},
            y label style={at={(axis description cs: 0,1.05)},above},
            x label style={at={(axis description cs: 1.05,0.5)},right},
            ymin=-4, ymax=4,
            xmin=0, xmax=8,
            ytick={-4,-2,...,4},
            xtick={0,2,...,8},
            minor y tick num=1,
            minor x tick num=1,
            grid=both,
            clip=false,
            ]
            \node[left] at (0,0) {0};
            \addplot[very thick,domain=0:8] {-2};
        \end{axis}
    \end{tikzpicture}
    \end{center}
\end{enumerate}

\clearpage

\section{Force Interactions}

\subsection*{Vocabulary}

\begin{itemize}[itemsep=-3pt,topsep=0pt]
    \item Define \textbf{\gls{system}} as \glsdesc{system}.
    \item Define \textbf{\gls{force}} as \glsdesc{force}.
    \item Define \textbf{\gls{net force}} ($F_\text{\scriptsize net}$) as \glsdesc{net force}.
    \item Define \textbf{\gls{external force}} as \glsdesc{external force}.
    \item Define \textbf{\gls{net external force}} ($F_\text{\scriptsize net}$) as \glsdesc{net external force}.
    \item Define \textbf{\gls{free-body diagram}} as \glsdesc{free-body diagram}.
    \item Define \textbf{\gls{Newton's first law of motion}} as stating that \glsdesc{Newton's first law of motion}.
    \item Define \textbf{\gls{inertia}} as \glsdesc{inertia}.
    \item Define \textbf{\gls{tension}} as \glsdesc{tension}.
    \item Define \textbf{\gls{friction}} as \glsdesc{friction}.
    \item Define \textbf{\gls{normal force}} as \glsdesc{normal force}.
    \item Define \textbf{\gls{gravity}} as \glsdesc{gravitational force} (later, \glsdesc{gravity}).
    \item Define \textbf{\gls{applied force}} as \glsdesc{applied force}.
    \item Define \textbf{\gls{mass}} as \glsdesc{mass}.
    \item Define \textbf{\gls{Newton's third law of motion}} as stating that \glsdesc{Newton's third law of motion}.
\end{itemize}

\subsection*{Fundamental Principles}



\begin{itemize}[itemsep=0pt,topsep=-3pt]
    \item \textbf{Know} that an object will not speed up, slow down, or change direction on its own, and that it takes a net external force to change motion.
    \item \textbf{Know} that two interacting objects exert equal and opposite forces on each other.
\end{itemize}

\subsection*{Mathematical Formulas}

\begin{itemize}[topsep=-3pt,itemsep=0pt]
    \item \textbf{Know} that the base SI unit of force is the newton (N).
    \item \textbf{Know} that the base SI unit of mass is the kilogram (kg).
    \item \textbf{Know} that a newton (N) is equivalent to a kilogram times a meter per second squared ($\mathrm{kg\cdot m/s^2}$).
\end{itemize}

\subsection*{Learning Goals}

\begin{enumerate}[itemsep=0pt]
    \item \textbf{Indicate} whether forces are balanced or unbalanced in a free-body diagram.
    \item \textbf{Indicate} whether forces are balanced or unbalanced from graphs of position, velocity, or acceleration vs.\,time.
    \item \textbf{Indicate} whether forces are balanced or unbalanced from a motion map.
    \item \textbf{Calculate} the vector sum of forces on a free-body diagram in one dimension, to find both the direction and magnitude of the net force vector.
    \item \textbf{Identify} when a tension force is exerted on an object.
    \item \textbf{Identify} the direction of the normal force for a box on an inclined surface.
\end{enumerate}

\subsection*{Assessment}

\begin{enumerate}
    \item Are the forces balanced or unbalanced?
    
    \begin{center}
    \begin{tikzpicture}[x=5mm,y=5mm]
        \draw[step=1,lightgray] (-3,-3) grid (3,3);
        \fill (0,0) circle (0.4);
        \draw[ultra thick,->] (0,0) -- (+2,0);
        \draw[ultra thick,->] (0,0) -- (-3,0);
        \draw[ultra thick,->] (0,0) -- (0,-2);
        \draw[ultra thick,->] (0,0) -- (0,+2);
    \end{tikzpicture}
    \end{center}

    \item In the figure below, are forces balanced or unbalanced?

    \begin{center}
        \begin{tikzpicture}
        \begin{axis}[width=5.2cm,height=4.2cm,
            axis lines=left,
            xlabel = {Time (s)},
            ylabel = {Position (m)},
            ymin=0, ymax=8,
            xmin=0, xmax=8,
            ytick={0,2,...,8},
            xtick={0,2,...,8},
            minor y tick num=1,
            minor x tick num=1,
            grid=both,
            ]
            \addplot[very thick,domain=0:8] {6/8*x};
            \end{axis}
        \end{tikzpicture}
    \end{center}

    \item Are forces balanced or unbalanced?

    \begin{center}
    \begin{tikzpicture}
        \begin{axis}[width=5.5cm,height=4.5cm,
            axis lines=center,
            xlabel = {Time (s)},
            ylabel = {Acceleration ($\mathrm{m/s^2}$)},
            y label style={at={(axis description cs: 0,1.05)},above},
            x label style={at={(axis description cs: 1.05,0.5)},right},
            ymin=-4, ymax=4,
            xmin=0, xmax=8,
            ytick={-4,-2,...,4},
            xtick={0,2,...,8},
            minor y tick num=1,
            minor x tick num=1,
            grid=both,
            clip=false,
            ]
            \node[left] at (0,0) {0};
            \addplot[very thick,domain=0:8] {3};
        \end{axis}
    \end{tikzpicture}
    \end{center}

    \item Are forces balanced or unbalanced?

    \begin{center}
    \begin{tikzpicture}
        \begin{axis}[width=5.4cm,height=4.5cm,
            axis lines=center,
            xlabel = {Time (s)},
            ylabel = {Velocity ($\mathrm{m/s}$)},
            y label style={at={(axis description cs: 0,1.05)},above},
            x label style={at={(axis description cs: 1.05,0.5)},right},
            ymin=-20, ymax=20,
            xmin=0, xmax=8,
            ytick={-20,-12,...,20},
            xtick={0,2,...,10},
            minor y tick num=1,
            minor x tick num=1,
            grid=both,
            clip=false,
            ]
            % \node[left] at (0,0) {0};
            \addplot[very thick,domain=0:8] {16 - 3*\x};
        \end{axis}
    \end{tikzpicture}
    \end{center}

    \item The motion map below depicts an object moving with negative velocity. Are the forces balance or unbalanced?

    \begin{center}
        \begin{tikzpicture}[y=15cm,x=15cm]
            \foreach \x in {0,0.1,...,1} {
                \fill (\x*\x,0) circle (1mm);
            }
            \draw[thick,->] (0,0.05) -- ++(0.07,0) node[right] {$+x$};
        \end{tikzpicture}
    \end{center}

    \item In the free-body diagram below, each square represents \SI{10}{N}. At $t = 0$, the object moves at a speed of \SI{6}{m/s}. If the external forces do not change, what is object's speed at $t = \SI{5}{s}$?
    
    \begin{center}
    \begin{tikzpicture}[x=5mm,y=5mm]
        \draw[step=1,lightgray] (-3,-3) grid (3,3);
        \fill (0,0) circle (0.4);
        \draw[ultra thick,->] (0,0) -- (+2,0);
        \draw[ultra thick,->] (0,0) -- (-2,0);
        \draw[ultra thick,->] (0,0) -- (0,-3);
        \draw[ultra thick,->] (0,0) -- (0,+3);
    \end{tikzpicture}
    \end{center}

    \item Calculate the net force in the free-body diagram below.

    \begin{center}
    \begin{tikzpicture}[x=3mm,y=3mm]
        \draw[very thick,->] (0,0) -- (12,0) node[right] {\SI{12}{N}};
        \draw[very thick,->] (0,0) -- (-8,0) node[left] {\SI{8}{N}};
        \fill (0,0) circle (2mm);
    \end{tikzpicture}
    \end{center}

    \item A box is suspended below an anchoring rod by two cables, as shown below. What type of force do the cables exert on the box?
    
    \begin{center}
    \begin{tikzpicture}[x=7mm,y=7mm]
        \draw[fill=black!10] (-0.5,-0.5) rectangle ++(1,1);
        \draw (0,0.5) -- ++({-2},{2*tan(30)});
        \draw (0,0.5) -- ++({+2},{2*tan(30)});
        \draw[pattern=north west lines, pattern color=black] (-2.5,1.68) rectangle ++(5,0.2);
    \end{tikzpicture}
    \end{center}

    \item The free-body diagram below shows the instant an object is moving at constant speed to the left across a frictionless surface. Some time later, the object a rough surface with plenty of friction. Sketch the free-body diagram when the object moves across the rough surface.

    \begin{center}
    \begin{tikzpicture}[x=4mm,y=4mm]
        \draw[step=1,lightgray] (-3,-3) grid (3,3);
        \fill (0,0) circle (0.4);
        \draw[ultra thick,->] (0,0) -- (0,-3);
        \draw[ultra thick,->] (0,0) -- (0,+3);
    \end{tikzpicture}
    \end{center}

    \item What is the direction of the \underline{normal} force on the box below?
    
    \begin{center}
    \begin{tikzpicture}[x=5mm,y=5mm,rotate=-40]
        \draw (-2.5,-1) -- (2.5,-1);
        % \fill (0,0) circle (2pt);
        \draw[thick,fill=black!10] (-1,-1) rectangle (1,1);
        % \draw[->,thick] (0,0) -- (0,{3*cos(40)}) node[above] {$F_\mathrm{N}$};
        \draw[rotate around={40:(2.5,-1)}] (-1,-1) node[above,shift={(5mm,0)}] {\ang{40}} -- (2.5,-1);
        % \draw[->,rotate=40] (0,0) -- (0,-3) node[below] {$F_g$};
    \end{tikzpicture}    
    \end{center}

    \item What is the direction of the \underline{gravity} force on the box below?
    
    \begin{center}
    \begin{tikzpicture}[x=5mm,y=5mm,rotate=-40]
        \draw (-2.5,-1) -- (2.5,-1);
        \draw[thick,fill=black!10] (-1,-1) rectangle (1,1);
        \draw[rotate around={40:(2.5,-1)}] (-1,-1) node[above,shift={(5mm,0)}] {\ang{40}} -- (2.5,-1);
    \end{tikzpicture}    
    \end{center}

    \item If mass measures the amount of matter in an object, can two objects that are the same size have different masses?

    \item At the instant shown below, objects 1 and 2 are undergoing a collision. The magnitude of the force on object 1 by object 2 is $F_{12}$, and the magnitude of the force on object 2 by object 1 is $F_{21}$. Object 2 has a greater mass than object 1. How do the forces $F_{12}$ and $F_{21}$ compare?
    
    \begin{center}
    \begin{tikzpicture}[y=3cm,x=3cm]
        \draw[fill=cyan!50] (1.84,0) circle (0.16) node {1}; 
        \draw[fill=red!50] (2.25,0) circle (0.25) node {2};
    \end{tikzpicture}
    \end{center}

    \item Figure 1 below shows a box that is held at rest by a rope of negligible mass attached to a pulley. Points A and B are where the rope connects to the box and pulley, respectively. Figure 2 shows the force on the box at point A. Draw the force on the pulley at point B.
    
    \begin{center}
    \begin{tikzpicture}
        \draw[fill=black!20] (0,0) circle (1) node[above=1] {pulley};
        \fill (0,0) circle (0.02);
        \draw (1,0) -- (1,-2.5);
        \begin{scope}[shift={(1,-2.5)},x=2.5mm,y=2.5mm]
            \draw[fill=black!10] (-1,-1) rectangle (1,1);
        \end{scope}
        \fill (1,0) circle (0.08) node[right] {B};
        \fill (1,-2.25) circle (0.08) node[shift={(0.2,0.25)}] {A};
        \node[below=5mm] at (current bounding box.south) {Figure 1};
    \end{tikzpicture}
    \hspace{1cm}
    \begin{tikzpicture}[x=5mm,y=5mm]
        \draw[step=1,lightgray] (-3,-3) grid (3,3);
        \node[below=5mm] at (current bounding box.south) {Figure 2};
        \draw[ultra thick,->] (0,-2) -- ++(0,4);
    \end{tikzpicture}
    \end{center}
\end{enumerate}

\clearpage

\section{Newton's Second Law and Universal Gravitation}

\subsection*{Vocabulary}

\begin{itemize}[itemsep=0pt]
    \item Define \textbf{\gls{Newton's second law of motion}} as stating that \glsdesc{Newton's second law of motion}.
    \item Define \textbf{\gls{Newton's universal law of gravitation}} as stating that \glsdesc{Newton's universal law of gravitation}.
    \item Define the \textbf{\gls{gravitational constant}} ($G$) as \glsdesc{gravitational constant}.
    \item Define \textbf{\gls{weight}} as \glsdesc{weight}.
\end{itemize}

\subsection*{Fundamental Principles}

\subsection*{Mathematical Formulas}

\begin{itemize}[itemsep=0pt]
    \item \textbf{Know} that Newton's second law of motion is give by $\displaystyle a = \frac{F_\mathrm{net}}{m}$ or equivalently as $F_\text{\scriptsize net} = ma$.
    \item \textbf{Know} that Newton's universal law of gravitation is given by the equation $\displaystyle |F_g| = \frac{G m_1 m_2}{r^2}$
    \item \textbf{Know} that the gravitational constant is $\displaystyle G = \SI{6.67e-11}{N\cdot m^2/s^2}$.
    \item \textbf{Know} that $r$ in the law of gravitation is the distance between the centers of mass.
\end{itemize}

\subsection*{Learning Goals}

\begin{enumerate}[itemsep=0pt]
    \item \textbf{Translate} forces on a free-body diagram to a net force equation in one dimension.
    \textbf{Calculate} acceleration and force using Newton's second law.
    \item \textbf{Calculate} the magnitude of the gravitational force between two objects, expressing the answer in scientific notation.
\end{enumerate}

\subsection*{Assessment}

\begin{enumerate}[itemsep=0pt]
    \item Write the net force equation for the forces in the horizontal direction.

    \begin{center}
    \begin{tikzpicture}[x=1.2cm,y=1.2cm]
        \draw[->,thick] (0,0) -- (1,0) node[right] {$F_\text{\scriptsize A}$};
        \draw[->,thick] (0,0) -- (-1,0) node[left] {$F_f$};
        \draw[->,thick] (0,0) -- (0,1) node[above] {$F_\text{\scriptsize N}$};
        \draw[->,thick] (0,0) -- (0,-1) node[below] {$F_g$};
        \fill (0,0) circle (4pt);
    \end{tikzpicture}
    \end{center}
\end{enumerate}

\clearpage

\section{One-Dimensional Kinematics}

\subsection*{Vocabulary}

\begin{itemize}[topsep=-3pt,itemsep=0pt]
    \item Define \textbf{\gls{kinematic equations}} as \glsdesc{kinematic equations}.
    \item Define \textbf{\gls{free fall}} as \glsdesc{free fall}.
    \item Define \textbf{\gls{hang time}} as \glsdesc{hang time}.
    \item Define \textbf{\gls{air resistance}} as \glsdesc{air resistance}.
    \item Define \textbf{\gls{negligible}} as \glsdesc{negligible}.
    \item Define \textbf{\gls{projectile}} as \glsdesc{projectile}.
\end{itemize}

\subsection*{Mathematical Formulas}

\begin{itemize}[topsep=-3pt,itemsep=0pt]
    \item \textbf{Know} velocity as a function of time: $\displaystyle v_f = v_i + at$
    \item \textbf{Know} position as a function of time: $\displaystyle \Delta x = v_i t + \frac{1}{2}at^2$
    \item \textbf{Know} velocity as a function of position: $\displaystyle v_f^2 = v_i^2 + 2a\Delta x$
    \item \textbf{Know} average velocity to find displacement: $\displaystyle \Delta x = \frac{1}{2}(v_i + v_f)t$
\end{itemize}

\subsection*{Learning Goals}

\begin{itemize}[itemsep=0pt]
    \item \textbf{Sketch} a graph of vertical velocity vs.\,time for the entire flight of an object projected vertically from the ground.
\end{itemize}


\subsection*{Assessment}

\clearpage

\section{Two-Dimensional Motion}

\subsection*{Vocabulary}

\vspace{-6pt}

\begin{itemize}[itemsep=0pt]
    \item Define \textbf{\gls{projectile motion}} as \glsdesc{projectile motion}.
    \item Define \textbf{\gls{trajectory}} as \glsdesc{trajectory}.
    \item Define \textbf{\gls{apex}} as \glsdesc{apex}.
    \item Define \textbf{\gls{horizontally launched projectile}} as \glsdesc{horizontally launched projectile}.
    \item Define \textbf{\gls{impact speed}} as \glsdesc{impact speed}.
    \item Define \textbf{\gls{circular motion}} as \glsdesc{circular motion}.
    \item Define \textbf{\gls{centripetal acceleration}} as \glsdesc{centripetal acceleration}.
    \item Define \textbf{\gls{centripetal force}} as \glsdesc{centripetal force}.
    \item Define \textbf{\gls{Newton's universal law of gravitation}} as a law stating that \glsdesc{Newton's universal law of gravitation}.
\end{itemize}

\subsection*{Fundamental Principles}

\begin{itemize}[itemsep=0pt]
    \item \textbf{Know} that, in projectile motion, the trajectory's shape is a parabola.
\end{itemize}

\subsection*{Mathematical Formulas}




\subsection*{Learning Goals}

\subsection*{Assessment}


\clearpage

\section{Energy}

\subsection*{Vocabulary}

\vspace{-6pt}

\begin{itemize}[itemsep=0pt]
    \item Define \textbf{\gls{kinetic energy}} as \glsdesc{kinetic energy}.
    \item Define \textbf{\gls{joule}} as \glsdesc{joule}.
    \item Define \textbf{\gls{work}} as \glsdesc{work}.
    \item Define \textbf{\gls{energy}} as \glsdesc{energy}.
    \item Define \textbf{\gls{potential energy}} as \glsdesc{potential energy}.
    \item Define \textbf{\gls{gravitational potential energy}} as \glsdesc{gravitational potential energy}.
    \item Define \textbf{\gls{conservation of energy}} as \glsdesc{conservation of energy}.
    \item Define \textbf{\gls{mechanical energy}} as \glsdesc{mechanical energy}.
    \item Define \textbf{\gls{conservation of mechanical energy}} as \glsdesc{conservation of mechanical energy}.
\end{itemize}

\subsection*{Fundamental Principles}

\subsection*{Mathematical Formulas}

\subsection*{Learning Goals}

\subsection*{Assessment}

\clearpage

\section{Momentum}

\subsection*{Vocabulary}

\vspace{-6pt}

\begin{itemize}[itemsep=0pt]
    \item Define \textbf{\gls{momentum}} as \glsdesc{momentum}.
    \item Define \textbf{\gls{momentum vs. time graph}} as \glsdesc{momentum vs. time graph}.
    \item Define \textbf{\gls{impulse}} as \glsdesc{impulse}.
    \item Define \textbf{\gls{impulse-momentum theorem}} as \glsdesc{impulse-momentum theorem}.
    \item Define \textbf{\gls{conservation of momentum}} as \glsdesc{conservation of momentum}.
    \item Define \textbf{\gls{elastic collision}} as \glsdesc{elastic collision}.
    \item Define \textbf{\gls{inelastic collision}} as \glsdesc{inelastic collision}.
    \item Define \textbf{\gls{perfectly inelastic collision}} as \glsdesc{perfectly inelastic collision}.
\end{itemize}

\subsection*{Fundamental Principles}

\subsection*{Mathematical Formulas}

\subsection*{Learning Goals}

\subsection*{Assessment}

\clearpage

\section{Oscillations and Mechanical Waves}

\clearpage

\section{Electrostatics and Circuits}

\clearpage

\section{Electromagnetism}

\clearpage

\section{Electromagnetic Waves and Quantum Physics}


\clearpage

\phantomsection
\addcontentsline{toc}{section}{Glossary}
\printnoidxglossaries



\end{document}